---
title: "Trash related service requests and ECB citations"
output: html_document
css: "style.css"
params:
  area_name: "Ellwood Park/Monument"
  area_type: "neighborhood"
  area_label: NULL
  year_start: 2016
  # Map parameters
  asp: "9:6.25" # "17:11"  # "3:2"  # "16:9"
  diag_ratio: 0.08
---

<!--Increases the size of the display after the document is Knitted-->
<style type="text/css"> 
.main-container {
  max-width: 80%;
  margin-left: auto;
  margin-right: auto;
}
</style>

::::::{id="body"}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      error = FALSE,
                      cache = FALSE)

library(sf)
library(mapbaltimore)
library(ggplot2)
library(dplyr)
library(stringr)
library(lubridate)
library(patchwork)
library(gt)

# Call scripts for get_area, get_requests, get_citations, and get_area_block
knitr::read_chunk("misc.R")

# Set default theme
prototype_theme <- theme_minimal() %+replace%
  theme(
    axis.line = element_line(color = "black"),
    axis.title = element_text(size = 13),
    axis.text = element_text(face = "bold"),
    axis.ticks = element_line(colour = "black", size = 1),
    legend.title = element_text(face = "bold", hjust = 0),
    legend.text = element_text(size = 11)
  )
  
theme_set(prototype_theme)
  
#Modify bar plots
update_geom_defaults("col", list(alpha = 0.6))
```

```{r get_area}
# Gets the area type, name, nearby areas, and area outline
```

```{r get_requests}
# Gets selected type of requests, 6-months prior, area request info, and nearby areas request info. 
```

```{r get_citations}
# Gets citations based on area type, provides citation source, area citations info, and nearby areas citations info.
```

```{r get_area_blocks}
# Gets blocks from the chosen area
```

::::{id="main_maps"}
```{r city_and_neighborhood_maps}
#Creates map of Baltimore City.
city_map <- ggplot() +
  geom_sf(data = baltimore_city, fill = "grey") +
  geom_sf(data = neighborhoods, fill = NA, color = "black") +
  geom_sf(data = area, fill = "red", color = "red") +
  geom_sf(data = baltimore_city, size = 1, color = "black", fill = NA) +
  ggtitle("Baltimore City") +
  theme_void()


#Creates map of neighborhood.
area_map <- ggplot() +
  layer_area_data(area = area, data = parks, fill = "forestgreen", alpha = 0.8) +
  geom_sf(data = blocks, fill = NA, color = "black") +        #Blocks layer
  geom_sf(data = area, color = "black", size = 1, fill = NA) +
  theme_void() +
  labs(
    title = params$area_name
  )

#displays maps
city_map + area_map
```
::::

<!-- Beginning of Requests div -->
:::::{id="service_requests"} 
# Service Requests

::::{id="description"}
The 311 Service Requests describe complains about trash in Baltimore City. 
The type of complains included in this analysis are Dirty Alleys, Dirty Streets, Illegal Dumping, and Sanitation Property. 

The location and the type of 311 service requests in the **`r params$area_name` `r params$area_type`** over the past **`r months_prior` months** is shown on the map below. 
::::

::::{class="row"}
:::{class="column left"}
```{r service_requests}
#Creates map of service requests colored by request type
area_requests %>% 
  ggplot() +
  layer_area_data(area = area, data = parks, fill = "forestgreen", alpha = 0.8) +
  geom_sf(data = blocks, fill = NA, color = "black") +        #Blocks layer
  geom_sf(data = area, color = "black", size = 1.25, fill = NA) +  #Neighborhood outline layer
  geom_sf(aes(color = sr_type)) +
  theme_void() +
  labs(
    title = glue::glue("Service Requests by type in {area_label}"),
    color = "Service Request Type"
  )

#Counts the total number of area requests
total_requests <- nrow(area_requests)
```
:::

:::{class="column right"}
:::{id="info"}
**`r total_requests`** Service Requests were made in `r area_label` over the past **`r months_prior` months**.
:::
:::
::::

::::{id="extra_info"}
```{r sr_request_types}
#Plots the service requests types in the chosen area
area_requests %>%
  st_drop_geometry() %>%
  count(sr_type, sort = TRUE) %>%
  ggplot() +
  geom_col(aes(x = reorder(sr_type, n), y = n, fill = sr_type), show.legend = FALSE) +
  xlab("Request Type") +
  ylab("Total Requests") +
  labs(
    title = glue::glue("Service Requests in {area_label}"),
    fill = "Request Type"
  )
```
::::

:::::
<!-- End of Requests div -->

<!-- Beginning of Citations div -->
:::::{id="ecb_citations"} 
# ECB Citations

::::{id="description"}
The ECB Citations measure the city's code enforcement activities. This analysis focuses on the following citations: 
Bulk Trash, Exterior Sanitary Maintenance, Occupant Trash Disposal, and Trash Accumulation.  

The location and the type of ECB Citations within the **`r area$name` `r params$area_type`** over the past **`r months_prior`** months is shown on the map below. 
::::

::::{class="row"}
:::{class="column left"}
```{r citations}
#Maps the area citations
area_citations %>% 
  ggplot() +
  layer_area_data(area = area, data = parks, fill = "forestgreen", alpha = 0.8) + 
  geom_sf(data = blocks, fill = NA, color = "black") +        #Blocks layer
  geom_sf(data = area, color = "black", size = 1.25, fill = NA) +  #Neighborhood outline layer
  geom_sf(aes(color = description)) +
  area_layer +
  theme_void() +
  labs(
    color = "Description",
    title = glue::glue("ECB Citations in {area_label}"),
    caption = citations_source
  )

#Counts the total number of area citations
total_citations <- nrow(area_citations)
```
:::

:::{class="column right"}
:::{id="info"}
**`r total_citations`** ECB Citations were made in `r area$name` over the past **`r months_prior` months**.
:::
:::
::::

::::{id="extra_info"}
```{r ecb_citations_types}
#Plots the citation types in the chosen area
# FIXME: How to modify the x-axis marks to make them look like the map's legend?

area_citations %>%
  st_drop_geometry() %>%
  count(description, sort = TRUE) %>%
  ggplot() +
  geom_col(aes(x = reorder(description, n), y = n, fill = description), show.legend = FALSE) +
  ggtitle("ECB Citations") +
  xlab("Citation Type") +
  ylab("Total Citations") +
  labs(
    title = glue::glue("ECB Citations in {area_label}"),
    fill = "Citation Type",
    caption = citations_source
  )
```
::::
:::::
<!-- End of Citations div -->

<!-- Beginning of Trash Issues div -->
:::::{id="trash_issues"}
# Trash Issues

::::{id="description"}
The area blocks with the highest number of citations since **`r params$year_start`** in the **`r area$name` `r params$area_type`** are the following. 
::::

::::{class="row"}
:::{class="column left"}
```{r trash_issues}
street_blocks %>% 
  ggplot() +
  layer_area_data(area = area, data = parks, fill = "forestgreen", alpha = 0.8) +
  geom_sf(data = area, color = "black", size = 1.25, fill = NA) +  #Neighborhood outline layer
  geom_sf(fill = NA, aes(color = street_block_number)) +
  geom_sf_label(aes(label = street_block_number), size = 1) +
  guides(color = "none") +
  theme_void() +
  labs(
    title = glue::glue("Numbered blocks in {area_label}")
  )

```
:::

:::{class="column right"}
:::{id="table"}
```{r block_table}
block_street_count <- citations %>%
  sf::st_drop_geometry() %>%
  count(street_name, block_number, name = "count", sort = TRUE) %>%
  relocate(block_number, .before = street_name)

block_street_count %>%
  slice_head(n = 10) %>%
  gt() %>%
  tab_header(
    title = glue::glue("10 blocks in {area_label} with most citations since {params$year_start}")
  ) %>%
  cols_label(
    block_number = "Block",
    street_name = "Street",
    count = "Citations"
  ) %>%
  tab_source_note(
    citations_source
  )
```
:::
:::
::::
:::::
<!-- End of Trash Issues div -->

<!-- Beginning of Nearby Areas div -->
:::::{id="nearby_areas"}
# Nearby Areas - Trash Issues

::::{id="description"}
The comparison between the number of service requests in the **`r area$name` `r params$area_type`** and **its surrounding 
neighborhoods** is shown on the table below.

This comparison is based on the number of service requests over the past **`r months_prior` months**.
::::

::::{id="comparison_table"}
```{r compare_with_nearby_areas}
# Rank of requests in area
request_comparison_table <- requests %>% 
  sf::st_drop_geometry() %>% 
  mutate(
    nearby = if_else(
      service_request_num %in% nearby_area_requests$service_request_num,
      "Nearby areas",
      as.character(knitr::combine_words(area$name))
      )
    ) %>% 
  select(nearby, sr_type, sr_status, days_to_close) %>% 
  gtsummary::tbl_summary(
    by = "nearby",
    label = list(
      sr_type = "Request type",
      sr_status = "Current status",
      days_to_close = "Days to close request"
    )
  ) %>% 
  gtsummary::as_gt() %>% 
  tab_source_note(
    source_note = glue::glue("Nearby areas include {knitr::combine_words(nearby_areas$name)}.")
  ) %>% 
  tab_header(
    title = glue::glue("Service requests in {knitr::combine_words(area$name)} and nearby areas")
  )

request_comparison_table

```
::::

::::{id="graphs"}
::::{class="row"}
:::{class="column left"}
```{r area_citations_over_time}
area_citations %>%
  ggplot() +
  geom_freqpoly(aes(x = violation_date, color = description)) +
  xlab("Date") +
  ylab("Total Citations") +
  labs(
    title = glue::glue("Citations Over Time in {area_label}"),
    color = "Citation Types"
  )
```
:::

:::{class="column right"}
```{r nearby_citations_over_time}
citations %>%
  ggplot() +
  geom_freqpoly(aes(x = violation_date, color = description)) +
  ggtitle("Citations Over Time in Nearby Areas") +
  xlab("Date") +
  ylab("Total Citations") +
  labs(
    color = "Citation Types"
  )
```
:::
::::
::::
:::::
<!-- End of Nearby Areas div -->

<!-- Beginning of areas' fewer citations, unpaid citations, sr requests div -->
:::::{id="fewer_issues"}
# Other Trash Information

::::{id="description"}
- Areas in the `r params$area_type` with the lowest number of citations since **`r params$year_start`**
- Map with the location of unpaid citations in **`r params$area_name`**
- Table of areas in the **`r params$area_name` `r params$area_type`** with the lowest number of citations
::::

::::{id="graphs"}
::::{class="row"}
:::{class="column left"}
```{r lowest_citations_table}
#Table of the top 10 areas with the least number of Citations in the area
block_street_count %>%
  slice_tail(n = 10) %>% #THIS LINE GETS THE BOTTOM VALUES OF THE TABLE
  gt() %>%
  tab_header(
    title = glue::glue("10 blocks in {area_label} with the least citations since {params$year_start}")
  ) %>%
  cols_label(
    block_number = "Block",
    street_name = "Street",
    count = "Citations"
  ) %>%
  tab_source_note(
    citations_source
  )
```
:::

:::{class="column right"}
```{r unpaid_citations_map}
area_citations %>% 
  filter(citation_status == "Open") %>% #Filters all other types of citation status except Open
  ggplot() +
  layer_area_data(area = area, data = parks, fill = "forestgreen", alpha = 0.8) +
  geom_sf(data = blocks, fill = NA, color = "black") +        #Blocks layer
  geom_sf(data = area, color = "black", size = 1.25, fill = NA) +  #Neighborhood outline layer
  geom_sf(aes(color = citation_status)) +
  theme_void() +
  labs(
    title = glue::glue("Unpaid Citations in {area_label}"),
    caption = citations_source,
    color = "Citation Status"
  )
```
:::
::::
::::
:::::
::::::



