---
title: "Trash related service requests and ECB citations"
output: html_document
params:
  area_name: ["Ellwood Park/Monument"]  #, "Baltimore Highlands"
  area_type: "neighborhood"
  year_start: 2016
  # If union is TRUE, the union_name is required
  union: FALSE # TRUE
  union_name: NULL
  # Map parameters
  asp: "9:6.25" # "17:11"  # "3:2"  # "16:9"
  diag_ratio: 0.08
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(sf)
library(mapbaltimore)
library(ggplot2)
library(dplyr)
library(stringr)
library(lubridate)
library(patchwork)
library(gt)

# Set default theme
theme_set(theme_minimal())
```

# Background

## Area

```{r get_area}
#Runs external script with the get_area code
source('get_area_script.R')

```

```{r map_area}
area_boundary_color <- "purple"

area_map <- ggplot() +
  layer_area_data(area = area, data = parks, fill = "forestgreen", alpha = 0.8) +
  layer_area_streets(area = area, color = "gray85") +
  layer_area_streets(area = area, sha_class = "all") +
  geom_sf(data = area, color = area_boundary_color, fill = NA) +
  theme_void() +
  labs(
    title = params$area_name
  )

city_map <- ggplot() +
  geom_sf(data = baltimore_city, fill = "white") +
  geom_sf(data = neighborhoods, fill = NA, color = "gray90", size = 0.4) +
  geom_sf(data = area, fill = area_boundary_color, color = area_boundary_color) +
  geom_sf(data = baltimore_city, size = 1.25, color = "black", fill = NA) +
  theme_void()

area_map +
  inset_element(city_map, 0.7, 0, 1, 0.3)
```

## Data on trash

- **311 service requests**  
  - The 311 service request dataset describes complains about trash in Baltimore City. 
Some of these complains include: Dirty Alleys, Dirty Streets, Illegal Dumping, and Sanitation Property.  
- **ECB citations**  
  - The ECB citations dataset measures the city's code enforcement activities. Particularly, we are 
focusing on the following citations: Bulk Trash, Exterior Sanitary Maintenace, Occupant Trash Disposal,
and Trash Accumulation.  

### Get selected requests

```{r get_requests}
#Runs external script with the get_requests code
source('get_requests_script.R')

```

### Get selected citations

```{r get_citations}
#Runs external script with the get_citations code
source('get_citations_script.R')

```

# Where is trash an issue in the area?

- Requests
  - Map of requests
- Citations
  - Citations by block
  - Citations by block number and street name

## Requests
  
  
Comparison between the 311 service requests in the `r area$name` neighborhood and its surrounding neighborhoods. 

```{r request_comparison_table}
# Rank of requests in area
request_comparison_table <- requests %>% 
  sf::st_drop_geometry() %>% 
  mutate(
    nearby = if_else(
      service_request_num %in% nearby_area_requests$service_request_num,
      "Nearby areas",
      as.character(knitr::combine_words(area$name))
      )
    ) %>% 
  select(nearby, sr_type, sr_status, outcome, days_to_close) %>% 
  gtsummary::tbl_summary(
    by = "nearby",
    label = list(
      sr_type = "Request type",
      sr_status = "Current status",
      outcome = "Request outcome", # TODO: Request outcome data is messy
      days_to_close = "Days to close request"
    )
  ) %>% 
  gtsummary::as_gt() %>% 
  tab_source_note(
    source_note = glue::glue("Nearby areas include {knitr::combine_words(nearby_areas$name)}.")
  ) %>% 
  tab_header(
    title = glue::glue("Service requests in {knitr::combine_words(area$name)} and nearby areas")
  )

request_comparison_table

#gt::gtsave(request_comparison_table, "request_comparison_table.png")

```
  
    
The location and the type of 311 service requests within the `r area$name` neighborhood
over the past `r months_prior` months is shown on the map below.

```{r map_requests_sr_type}
requests %>% 
  ggplot() +
  geom_sf(aes(color = sr_type)) +
  geom_sf(data = area, fill = NA, color = "black", linetype = "dashed") +
  theme_void() +
  labs(
    title = "Area requests by type",
    color = "Service Request Type"
  ) +
 # scale_color_viridis_d(direction = -1) +
  #scale_color_brewer(palette = "Dark2") +
  set_map_limits(area = area, dist = 50)

```
  
The time it takes to close each service request within the `r area$name` neighborhood
is shown below. 

```{r map_requests_days_to_close}
requests %>% 
  ggplot() +
  geom_sf(aes(color = days_to_close), alpha = 0.6) +
  area_layer +
  labs(
    color = "Days to close"
  ) +
  scale_color_viridis_c() +
  theme_void() +
  set_map_limits(area = area, dist = 50)

```

## Citations
  
Comparison between the citation status of `r area$name` and its surrounding neighborhoods.

```{r plot_citations_by_status_nearby}
citations %>%
  st_drop_geometry() %>% 
  count(nearby, citation_status, sort = TRUE) %>% 
  ggplot() +
  geom_col(aes(x = reorder(citation_status, n), y = n, fill = citation_status)) +
  coord_flip() +
  facet_wrap(~nearby) +
  xlab("Citation Status") +
  ylab("Total Citations") +
  labs(
    fill = "Citation Status"
  ) +
  ggtitle("Citation Status")

```
  
  
The location and the description of citations within `r area$name` over the past `r months_prior` is shown below.

```{r map_citations}
citations %>% 
  ggplot() +
  geom_sf(aes(color = description), alpha = 0.75) +
  area_layer +
  theme_void() +
  labs(
    color = "Description",
    title = "Ellwood Park and its surrounding neighborhoods"
  ) +
  set_map_limits(area = area, dist = 50)

```

## Citations by block

```{r get_area_blocks}
area_property <-
  get_area_property(area = area, crop = FALSE)

blocks <- area_property %>% 
  group_by(block) %>% 
  summarise(
    geometry = st_union(geometry)
  )


blocks %>% 
  ggplot() +
  geom_sf(fill = NA, color = "black") +
  geom_sf_label(aes(label = block)) +
  theme_void() +
  labs(
    title = "Blocks in Ellwood Park/Monument"
  )

 mapview::mapview(blocks)
 
```

```{r citation_block}
#Plots citations per block in descending order

area_citations %>%
  st_drop_geometry() %>%
  count(block, sort = TRUE) %>%
  ggplot() +
  geom_col(aes(x = reorder(block, n), y = n)) +
  coord_flip() +
  theme(axis.text.y = element_text(size = 5)) +
  ggtitle("Citations per Block") +
  xlab("Block") +
  ylab("Total Citations")
  
```

```{r citation_type}
area_citations %>%
  st_drop_geometry() %>%
  count(block, description, sort = TRUE) %>%
  ggplot() +
  geom_col(aes(x = reorder(block, n), y = n,  fill = description)) +
  coord_flip() +
  theme(axis.text.y = element_text(size = 5)) +
  ggtitle("Citation Types per Block") +
  xlab("Block") +
  ylab("Total Citations") +
  labs(caption = citations_source,
       fill = "Description"
       )

```

```{r citations_by_block_table}
area_citations %>%
  st_drop_geometry() %>%
  count(block, name = "citations", sort = TRUE) %>%
  gt()
```

### Plot citations by block number and street

```{r plot_block_number_street}
# FIXME: Figure out a reproducible way for distinguishing between groups of streets
ns_street_cutoff <- 500 # Block number equal or less than this are in the NS group
ew_street_cutoff <- 500 # Block numbers greater than this are in the EW group

block_street_count <- citations |>
  sf::st_drop_geometry() |>
  count(street_name, block_number, name = "count", sort = TRUE) |>
  relocate(block_number, .before = street_name)

plot_block_street <- function(df, highlight_n = 10) {
  df |>
  ggplot() +
  geom_point(aes(x = count, y = block_number), size = 3, alpha = 0.85, color = "purple") +
  facet_wrap(~ street_name) +
  gghighlight::gghighlight(count > highlight_n)
}

ns_streets <-
  block_street_count |>
  filter(block_number <= ns_street_cutoff) |> # Limit plot to N-S streets
  plot_block_street()

ew_streets <-
  block_street_count |>
  filter(block_number > ew_street_cutoff) |> # Limit plot to E-W streets
  plot_block_street()

ns_streets + ew_streets
```

```{r table_block_number_street}
block_street_count |>
  slice_head(n = 10) |>
  gt() |>
  tab_header(
    title = glue::glue("10 blocks in {knitr::combine_words(params$area_name)} with most citations since {params$year_start}")
  ) |>
  cols_label(
    block_number = "Block",
    street_name = "Street",
    count = "Citations"
  ) |>
  tab_source_note(
    citations_source
  )
```

# How are citations addressing the issue of trash in this area?

- Citation status
  - Citation status map
    -Does the map show any cluster(s) where citations are voided or appealed?
- Citations over time
- Total paid per citation map
  - Should the average amount paid per type of citation be found?

### Citation status 

```{r citations_status}
#plots the total number of citations per status
area_citations %>%
  ggplot() +
  geom_bar(aes(x = citation_status, fill = description)) +
  ggtitle("Citation Status") +
  ylab("Total Citations") +
  xlab("Citation Status") +
  labs(fill = "Description")
```

```{r citation_status_table}
area_citations %>%
  st_drop_geometry() %>%
  count(citation_status, name = "citations") %>%
  gt() %>%
  tab_header(
    title = glue::glue("Citation status in {knitr::combine_words(params$area_name)}")
  ) %>%
  cols_label(
    citation_status = "Citation Status",
    citations = "Citations"          
  )
```

```{r citation_status_by_block}
#plots the number of citations per block while breaking them down into their status
area_citations %>%
  st_drop_geometry() %>%
  count(block, citation_status, sort = TRUE) %>%
  ggplot() +
  geom_col(aes(x = reorder(block, n), y = n,  fill = citation_status)) +
  coord_flip() +
  theme(axis.text.y = element_text(size = 5)) +
  ggtitle("Citation Status per Block") +
  xlab("Block") +
  ylab("Citation Status") +
  labs(
    fill = "Citation Status"
  )

```

```{r citation_status_map}
#creates a map that shows the location of each citation based on their status
area_citations %>% 
  ggplot() +
  geom_sf(aes(color = citation_status)) +
  geom_sf(data = area, fill = NA, color = "black", linetype = "dashed") +
  theme_void() +
  labs(
    caption = citations_source,
    color = "Citation Status",
    title = "Citation Status "
  )

```

```{r citation_table}

citations %>% 
  sf::st_drop_geometry() %>%
  select(nearby, description, citation_status) %>% 
  gtsummary::tbl_summary(
    by = "nearby",
    label = list(
      citation_status = "Citation Status",
      description = "Description"
    )
  ) %>% 
  gtsummary::as_gt() %>% 
  tab_source_note(
    source_note = glue::glue("Nearby areas include {knitr::combine_words(nearby_areas$name)}.")
  ) %>% 
  tab_header(
    title = glue::glue("ECB Citations in {knitr::combine_words(area$name)} and nearby areas")
  )


#Checks how many times the typo in "accumulation" occurs in order to fix it. 

# citations %>%
#   filter(description == "TRASH ACCUMULATION") %>%   
#   nrow()
# 
# citations %>%
#   filter(description == "TRASH ACCUMULATIION") %>%     Typo occurs once (1)
#   nrow()

```
### Citations over time

```{r total_citations_over_time}
#review this graph (binwidth)
citations_over_time <- citations %>%
  ggplot() +
  geom_freqpoly(aes(x = violation_date)) +
 # labs(caption = citations_source) +
  ggtitle("Citations over Time") +
  xlab("Date") + 
  ylab("Total Citations")

plotly::ggplotly(citations_over_time)
```

```{r citation_status_over_time}
# review this graph - Different way to plot this out? Maybe 4 line graphs?
area_citations %>% 
  ggplot() +
  geom_freqpoly(aes(x = violation_date, color = citation_status)) +
  labs(caption = citations_source) +
  ggtitle("Citation Status over Time") +
  xlab("Date") + 
  ylab("Total Citations") +
  labs(color = "Citation Status")

```

### Total paid per citation

```{r total_paid_map}
#maps out the total paid per citation
area_citations %>%
  ggplot() +
  geom_sf(aes(color = total_paid)) +
  geom_sf(data = area, fill = NA, color = "black", linetype = "dashed") +
  theme_void() +
  scale_color_viridis_c() +
  labs(
    caption = citations_source,
    color = "Total Paid"
    )
```
# How area service requests addressing the issue of trash in this area?

-Service requests
  - What is the avg number of days it takes to close each type of request?
-Service requests over time
-Service requests status

```{r plot_requests}
#Change tittle? 
#plots the number of days it takes for each service request to be closed
area_requests %>% 
  ggplot(aes(x = days_to_close, y = sr_type, color = sr_type), alpha = 0.6) +
  geom_point() +
  theme_minimal() + 
  ggtitle("Service Requests Closure Time") +
  xlab("Days to Close") + 
  ylab("Request Type") +
  labs(
    color = "Service Request Type"
  )
```

```{r request_table}
#creates a table to summarize the total number of each type of service request
area_requests %>%
  st_drop_geometry() %>% 
  count(sr_type, name = "requests", sort = TRUE) %>% 
  gt() %>%
  tab_header(
    title = glue::glue("Service Requests in {knitr::combine_words(params$area_name)}")
  ) %>%
  cols_label(
    sr_type = "Service Request Type",
    requests = "Requests"          
  )
```

```{r request_over_time}
#plots the number of requests that were created per (day?) for each type of request
area_requests %>%
  ggplot() +
  geom_freqpoly(aes(x = created_date, color = sr_type)) +
  ggtitle("Requests over Time") +
  xlab("Date") +
  ylab("Total Requests") +
  labs(
    color = "Service Request Type"
  )
```

```{r request_status_table}
#creates a table of the status for all requests
area_requests %>%
  st_drop_geometry() %>% 
  count(sr_status, name = "requests", sort = TRUE) %>% 
  gt() %>%
  tab_header(
    title = glue::glue("Service Request Status in {knitr::combine_words(params$area_name)}")
  ) %>%
  cols_label(
    sr_status = "Service Request Status",
    requests = "Requests"          
  )

```

